Backscatter Modulation Design for Symbiotic Radio Networks

### 2021-09-03 main.tex
:pencil: Update signal model

	\begin{section}{System Model}
		\begin{figure}[!t]
			\centering
			\def\svgwidth{0.9\columnwidth}
			\import{assets/}{symbiotic_radio.pdf_tex}
			\caption{A single-user single-tag symbiotic radio system.}
			\label{fi:symbiotic_radio}
		\end{figure}
		As shown in Fig.~\ref{fi:symbiotic_radio}, we propose a single-user (UE) single-tag (TG) symbiotic radio network where the RF signal generated by the single-antenna Access Point (AP) is shared by two coexisting systems. In the primary AP-UE downlink system, the AP transmits to the single-antenna user. In the secondary AP-TG-UE backscatter system, the AP acts as the carrier emitter, the user serves as the backscatter reader, and the single-antenna tag modulates its information over the reradiated RF signal by varying the reflection coefficient. Denote the AP-UE direct channel as $h_{\mathrm{D}}$, the AP-TG forward channel as $h_{\mathrm{F}}$, and the TG-UE backward channel as $h_{\mathrm{B}}$. We consider the quasi-static block fading model and assume the CSI of the direct channel and the cascaded forward-backward channel $h_{\mathrm{C}} \triangleq h_{\mathrm{B}} h_{\mathrm{F}}$ are known at the AP.\footnote{Due to the lack of RF chains at the passive tag, accurate and efficient CSI acquisition at the AP can be challenging. One possible approach is that the AP sends known pilots, the tag responds in a pre-defined manner, and the user performs least-square estimation then feeds back to the AP \cite{Bharadia2015,Yang2015b,Guo2019e}.} It is assumed that the primary symbol $s$ follows standard CSCG distribution $\mathcal{CN}(0,1)$ and the secondary symbol $c$ employs $M$-PSK modulation by \eqref{eq:backscatter_modulation}. To provide a preliminary insight, we consider a parasitic symbiotic radio \cite{Long2020a} where the primary and secondary symbol periods are equal. \footnote{However, parasitic symbiotic radio requires fast tag-state switching and frequent synchronization, which can be challenging for passive tags.} The user simultaneously captures the signal from both primary and secondary links as\footnote{We assume the time difference of arrival from the AP-UE path and the AP-TG-UE path are negligible compared to the symbol period \cite{Guo2019b,Liang2020,Long2020a}.}
		\begin{equation}
			y = \sqrt{P} h_{\mathrm{D}} s + \sqrt{\alpha P} h_{\mathrm{C}} s c + n,
		\end{equation}
		where $P$ is the average transmit power of the AP and $n \sim \mathcal{CN}(0,1)$ is the additive white Gaussian noise.

		\begin{remark}
			The symbiotic radio network can be regarded as as a special case of Multiple Access Channel (MAC) because the AP and the tag simultaneously transmit to the user. It is known that Superposition Coding-Successive Interference Cancellation (SC-SIC) with different decoding orders can achieve different vertices of the MAC capacity region \cite{Goldsmith2005}. Therefore, most relevant papers proposed the user to decode the primary message first (by treating the tag interference as noise), cancel out its contribution from the received signal, then decode the secondary message. Since the direct channel is typically much stronger than the cascaded channel \cite{Ozdogan2020}, the primary decoding is expected to enjoy a high Signal-to-Interference-and-Noise Ratio (SINR) and the secondary decoding is ideally interference-free.
		\end{remark}

		To investigate how backscatter modulation potentially benefits the primary transmission, we instead start decoding from the secondary link. Once the tag message is successfully recovered, by combining the backscattered symbol with the cascaded channel, the uncertainty of the AP-TG-UE path can be eliminated, and its contribution to primary transmission can be modeled within the equivalent AP-UE channel. This is reminiscent of IRS-aided point-to-point transmission. Overall, the passive tag not only embeds its own message in the reflection pattern, but also influences the legacy transmission.

		\begin{remark}
			The main difference between symbiotic radio and conventional MAC is that the primary message also reaches the user from the backscatter link. This characteristic inspires one to decode the tag message first and model its contribution within the equivalent channel (i.e., unify backscatter decoding and channel training), instead of performing SIC. In such case, the primary transmission is able to achieve ergodic capacity with artificial channel variation created by the backscatter modulation.
		\end{remark}

>>>>

	\begin{section}{System Model}
		\begin{figure}[!t]
			\centering
			\def\svgwidth{0.9\columnwidth}
			\import{assets/}{symbiotic_radio.pdf_tex}
			\caption{A single-user single-tag symbiotic radio system.}
			\label{fi:symbiotic_radio}
		\end{figure}
		As shown in Fig.~\ref{fi:symbiotic_radio}, we propose a single-user (UE) single-tag (TG) symbiotic radio network where the RF signal generated by the single-antenna Access Point (AP) is shared by two coexisting systems. In the primary AP-UE downlink system, the AP transmits to the single-antenna user. In the secondary AP-TG-UE backscatter system, the AP acts as the carrier emitter, the user serves as the backscatter reader, and the single-antenna tag modulates its information over the reradiated RF signal by varying the reflection coefficient. Denote the AP-UE direct channel as $h_{\mathrm{D}}$, the AP-TG forward channel as $h_{\mathrm{F}}$, and the TG-UE backward channel as $h_{\mathrm{B}}$. We consider the quasi-static block fading model and assume the CSI of the direct channel and the cascaded forward-backward channel $h_{\mathrm{C}} \triangleq h_{\mathrm{B}} h_{\mathrm{F}}$ are known at the AP.\footnote{Due to the lack of RF chains at the passive tag, accurate and efficient CSI acquisition at the AP can be challenging. One possible approach is that the AP sends known pilots, the tag responds in a pre-defined manner, and the user performs least-square estimation then feeds back to the AP \cite{Bharadia2015,Yang2015b,Guo2019e}.} It is assumed that the primary symbol $s$ follows standard CSCG distribution $\mathcal{CN}(0,1)$ and the secondary symbol $c$ employs $M$-PSK modulation by \eqref{eq:backscatter_modulation}. To provide a preliminary insight, we consider a parasitic symbiotic radio \cite{Long2020a} where the primary and secondary symbol periods are equal. \footnote{However, parasitic symbiotic radio requires fast tag-state switching and frequent synchronization, which can be challenging for passive tags.} The user simultaneously captures the signal from both primary and secondary links as\footnote{We assume the time difference of arrival from the AP-UE path and the AP-TG-UE path are negligible compared to the symbol period \cite{Guo2019b,Liang2020,Long2020a}.}
		\begin{equation}
			y = \sqrt{p} h_{\mathrm{D}} s + \sqrt{\alpha p} h_{\mathrm{C}} s c + n,
		\end{equation}
		where $p$ is the average transmit power at the AP and $n \sim \mathcal{CN}(0,1)$ is the additive white Gaussian noise.

		\begin{remark}
			The symbiotic radio network can be regarded as as a special case of Multiple Access Channel (MAC) because the AP and the tag simultaneously transmit to the user. It is known that Superposition Coding-Successive Interference Cancellation (SC-SIC) with different decoding orders can achieve different vertices of the MAC capacity region \cite{Goldsmith2005}. Therefore, most relevant papers proposed the user to first decode the primary message (by treating the tag interference as noise), cancel out its contribution from the received signal, then decode the secondary message. Since the direct channel is typically much stronger than the cascaded channel \cite{Ozdogan2020}, the primary decoding is expected to enjoy a high Signal-to-Interference-and-Noise Ratio (SINR) and the secondary decoding is ideally interference-free.
		\end{remark}

		\begin{remark}
			The main difference between symbiotic radio and conventional MAC is that the primary message also reaches the user from the backscatter link. This characteristic inspires one to first decode the tag message, then model its contribution within the equivalent channel during primary decoding (i.e., unify secondary decoding and backscatter channel training), instead of performing SIC. In such case, for each fading block, the primary transmission is able to achieve ergodic capacity with artificial channel variation created by the backscatter modulation.
		\end{remark}

		To investigate how backscatter modulation potentially benefits the primary transmission, we start the decoding process from the secondary link. Since $\mathbb{E}[\lvert s \rvert^2] = \mathbb{E}[\lvert c \rvert^2] = 1$, by treating the primary interference as noise, the SINR for the user to decode tag message $c$ is
		\begin{equation}
			\gamma_c = \frac{\alpha p \lvert h_{\mathrm{C}} \rvert^2}{p \lvert h_{\mathrm{D}} \rvert^2 + 1}.
		\end{equation}
		Once the tag message is successfully recovered, we combine the backscattered symbol with the cascaded channel to eliminate the uncertainty of the AP-TG-UE path, and the received signal is essentially
		\begin{equation}
			y = \sqrt{p} (h_{\mathrm{D}} + \sqrt{\alpha} h_{\mathrm{C}} c) s + n \triangleq \sqrt{p} h(c) s + n,
		\end{equation}
		which is reminiscent of IRS-aided point-to-point transmission with the equivalent channel defined as
		\begin{equation}
			h(c) \triangleq h_{\mathrm{D}} + \sqrt{\alpha} h_{\mathrm{C}} c.
		\end{equation}
		That is to say, the passive tag not only embeds its own message in the reflection pattern, but also influences the legacy transmission. For primary transmission, backscatter modulation creates a fast fading channel within the coherence time, and the equivalent CSI is known at the receiver once the backscattered symbol is successfully decoded. Therefore, the ergodic capacity of primary transmission within each fading block is \cite{Tse2005}
		\begin{equation}
			R_s = \mathbb{E}_c \bigl[ \log_2 ( 1 + \lvert h(c) \rvert^2 ) \bigr].
		\end{equation}
		We aims to optimize the input distribution of $c$ (i.e., the p.m.f. of $M$-PSK) to achieve a flexible balance between the primary and secondary links.
